---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mstATA

<!-- badges: start -->
<!-- badges: end -->

The goal of mstATA is to ...

## Installation

You can install the development version of **mstATA** from [GitHub](https://github.com/) using `pak`:

```{r}
# install.packages("pak")
pak::pak("Hongchen030/mstATA")
```

Alternatively, you can use `remotes`:

```{r}
# install.packages("remotes")
remotes::install_github("Hongchen030/mstATA")
```

`mstATA` supports multiple mixed-integer programming solvers, including:

Gurobi – requires separate installation and a valid license.
Installation instructions are available here:
https://cran.r-project.org/web/packages/prioritizr/vignettes/gurobi_installation_guide.html
Gurobi requires a valid license. However, free academic licenses are available for qualified students and researchers through the Gurobi Academic Program.

HiGHS – available on CRAN via install.packages("highs")

GLPK (via Rglpk) – available on CRAN via install.package("Rglpk")

SYMPHONY (via Rsymphony) – available on CRAN via install.package("Rsymphony")

lpSolve (via lpSolveAPI) – available on CRAN via install.package("lpSolveAPI")


## Vignettes

The package includes several vignettes:

- **Getting Started with mstATA** – complete end-to-end workflow.
- **MST Assembly Examples: Bottom-Up, Top-Down, and Hybrid Strategies** – bottom-up, top-down, and extended-hybrid strategies.
- **Stimulus-Based Assessment in mstATA: Conditional Item Selection** – conditional item selection with stimulus constraints.
- **Formulation for Multiple Objectives** – multi-objective optimization strategies and evaluation.
- **Detecting and Resolving Infeasibility** - how to deal with infeasibility.

To browse available vignettes:

```{r}
browseVignettes("mstATA")
```

## Citation

If you use `mstATA` in research, please cite:

Chen, H. (2026). mstATA: An R Package for IRT-Based Multistage Test Assembly.
